Usability Criteria for Graphical Code Completion

### 2012-03-18 vlhcc11/survey-report.tex
Initial paper move.

\usepackage{enumerate}
\usepackage{rotating}
\usepackage{url}

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Active Code Completion}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Cyrus Omar, YoungSeok Yoon, Thomas D. LaToza, Brad A. Myers}
\IEEEauthorblockA{School of Computer Science\\
Carnegie Mellon University\\
Pittsburgh, PA, USA\\
\{comar,youngseok,tlatoza,bam\}@cs.cmu.edu}
\texttt{\url{https://github.com/cyrus-/graphite}}
}

@@

% make the title area
% \maketitle

% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
% \IEEEpeerreviewmaketitle

\section{Developer Survey}

Prior to implementation, we conducted a large online survey of professional software developers in order to extract criteria that govern whether our system would be useful to them. More specifically, we sought to address the following questions:

\begin{itemize}
\item What are specific use cases for active code completion in a professional setting? 
\item What are some functional criteria that predict whether a class would benefit from active code completion?
\item What are some usability criteria that should govern palette interface designs?
\item Which capabilities must a active code completion architecture support to enable these use cases and designs?
\end{itemize}

% In order to convey the active code completion concept clearly, we came up with three different use cases where we belive the active code completion technique will benefit developers frequently. For each use case, we designed three mock-up screenshots and showed them in sequence: before invoking the palette, during the interaction with the palette, and after the palette is closed and code is generated. By doing this, the respondents could understand the concept precisely and give useful feedback relevant to the above questions.

We posted a link to our survey on a popular  collaborative filtering website \cite{mehta_robust_2007} called {\tt /r/programming} on reddit.com\footnote{\texttt{http://www.reddit.com/r/programming}}. At the time of our study, the website listed approximately 340,000, though members were not required to registered to read the content,  and contained content that could be considered relevant to our intended audience. An additional 22 participants were recruited using a mass email to local computer science graduate students. The recruitment materials in both cases stated that we were seeking developers ``familiar with an object-oriented programming language like Java, C\# or Visual Basic and an integrated development environment like Eclipse or Visual Studio''.

Participants were asked to take an online survey that would take approximately 20 minutes to complete. Of the 696 people who started the survey by answering the first question, 473 participants completed it. Of these, 16 participants were computer science graduate students at Carnegie Mellon and 457 were from \texttt{/r/programming}.

Their responses revealed a number of interesting use cases and non-trivial design constraints for active code completion systems. Many of these findings can also be generalized to constrain the design of visual programming languages, particularly those that aim be useful to professional developers, rather than novice programmers (e.g., Barista \cite{ko_barista:_2006}). We summarize the survey questions and the results in the following subsections.
% In fact, active code completion can be considered a hybrid approach that borrows interaction techniques from visual languages while remaining compatible with conventional representations of program logic. This may help address some of the usability challenges previously associated with visual languages (cf. [?]).

\subsection{Familiarity with Programming Languages and Editors}

We first asked their level of familiarity with several programming languages, on a five-point Likert scale ranging from ``None'' to ``Expert''. 61.1\% of the respondents expressed themselves as experts on at least one language, and 35.7\% were very familiar with at least one language. On average, participants rated themselves as very familiar with Java, C, C++ and JavaScript, familiar with C\#, Python and PHP and somewhat familiar with Visual Basic and Perl.

We also asked participants to select the integrated development environments (IDEs) and code editors that they were familiar with. The Eclipse IDE was familiar to 87.1\% of participants. This was followed by Visual Studio at 66.0\%, Vi/Vim at 53.7\%, Netbeans at 37.7\%, Emacs at 24.8\% and IntelliJ IDEA at 16.4\%. Participants could also enter ``other'' choices. A number of editors and IDEs were entered, including Xcode, Textmate and Notepad++. None of these were dominant.

\subsection{Mock-ups}

\begin{figure}
  \includegraphics[width=\linewidth]{mockup-palette-small.png}
  \caption{An example mock-up palette for regular expressions shown to the survey participants. In the survey, they were shown the surrounding Eclipse IDE together.}
  \label{regex}
\end{figure}

Next, we presented participants with a series of mock-up palettes for a Color class, a regular expression class (Fig. \ref{regex}), and a SQL query class. Participants were also shown a mock-up demonstrating how a user could invoke the palette, and a mock-up showing the code that would be inserted once a selection had been made.

Before presenting each mock-up, we gathered information about the user's familiarity with the class and the strategies that they would likely use to instantiate the class under normal circumstances. For the Color class, the majority of participants indicated that they would look in the code completions menu (58.4\%) or in the class documentation (19.0\%) for a predefined constant. Another 14.0\% indicated that they would use an external tool to determine the RGB values. Few participants were unfamiliar with regular expressions and no participants were unfamiliar with SQL queries which provides further evidence that our participants were not novice developers. Fig. \ref{strategies} summarizes their usual strategies for instantiating regular expressions and SQL queries.

\begin{figure}
\begin{tabular}{lccc}
 & Regular Expressions &  & SQL\\
 \hline
Separate test script & 29.6\% &  & 15.4\%\\
Guess and check & 14.0\% &  & 16.1\%\\
External tool & \textbf{37.9\%} &  & \textbf{58.6\%}\\
Search for examples & 12.3\% & & 5.1\% \\
Other & 6.2\% & & 4.9\% \\
\hline
\end{tabular}
\caption{Typical strategies for regular expressions and SQL queries.}
\label{strategies}
\end{figure}

\begin{figure}
\begin{tabular}{crccccc}\\\\
\textsc{class}
& 
& \begin{rotate}{20}Nearly every time\end{rotate}
& \begin{rotate}{20}Most of the time\end{rotate}
& \begin{rotate}{20}Some of the time\end{rotate}
& \begin{rotate}{20}Rarely\end{rotate}
& \begin{turn}{20}Never\end{turn}\\
\hline
Color &\vline& 9.6\% & 22.1\% & \textbf{32.4\%} & 28.2\% & 7.7\%\\
RegExp &\vline& \textbf{36.6\%} & 29.5\% & 21.8\% & 7.3\% & 4.8\%\\
SQL & \vline &18.2\% & 19.3\% & \textbf{30.9\%} & 20.4\% & 11.4\%\\
\hline
\end{tabular}
\caption{The distribution of responses to the question: ``Consider situations where you need to instantiate the [specified] class. What portion of the time, in these situations, do you think you would use this feature?''}
\label{frequencies}
\end{figure}

Participants were positive about our mock-ups. We asked the users to rate the usefulness of each mock-up palette, and their responses are summarized in Fig. \ref{frequencies}. For each palette, more than half of the respondents expressed that they would use it at least some of the time. Especially, people liked the regular expression palette a lot.

\section{Design Implications}

We analyzed the open-ended responses to extract design principles and constraints. We got 193 responses for the color palette, 129 for the regular expression palette, and 142 for the SQL palette. Also, we got total 119 suggestions about what other types of classes could benefit from similar types of palettes. The extracted principles and constraints are summarized in the following subsections. The parenthesized number following each design constraint indicates how many people commented on that issue.

@@

System design constraints are the criteria that should be taken into account in the design of the overall system architecture.

\subsubsection{Handling Separation of Concerns (183?)}

Participants were very wary of including constant data into the source code directly. Many responses expressed general sentiment or specific preference for project-wide color theme or external resource file. Sometimes they explicitly suggested to add capability to handle the external resource files directly from the palette. Interestingly, most people seemed to consider regular expressions as part of the program logic. There were few people who complained about including the regular expression pattern string in the source code as opposed to the other two classes.

%Examples of feedbacks in this category include:
%
%\begin{itemize}
%	\item The use of magic numbers in the source code is a bad design (54)
%	\item Color picking is the `designer's job' (33)
%	\item Would only use to define named constants (17)
%	\item Would rather use color scheme / theme (3)
%	\item Would use Object-Relational Mapping (ORM) / Language INtegrated Query (LINQ) instead of raw SQL query string (27 of 142)
%	\item These palettes might be useful for prototyping, but not for real production (14)
%	\item Instead of including the test cases into comments, they should be stored as unit tests (35 of 143)
%\end{itemize}

\subsubsection{Reversibility (19?)}

Some participants expressed that it should be able to bring the palettes back when they want to modify the parameters after they used the palettes.

\subsubsection{Palette Settings and State (41?)}

Many participants wanted to be able to configure the palettes or have the palettes maintain state even after the palettes are hidden. % The followings are the comments that are related to this category.

%\begin{itemize}
%	\item Want recently used regular expressions, and control over the comments (12)
%	\item Persistent database connection information is needed / want to manage connection pools (9)
%	\item Want recent colors, capability of setting favorite colors (20)
%\end{itemize}

\subsubsection{Interaction with Code Context (13?)}

Some people wanted to interact with the code context. For example, people expressed that it would be great if the SQL query palette was capable of dynamically creating SQL query string by combining multiple string variables, or by assigning variables for the parameters.

\subsubsection{Performance (?)}

@@

\subsubsection{IDE Independence (?)}

@@

\subsection{General Palette Design Considerations}

General palette design considerations are about how to make more usable palettes.

\subsubsection{Simplicity vs. Capability}

@@

\subsubsection{Keyboard Navigability (12)}

@@

\subsection{Suggestions from the Participants}
At the end of the survey, we solicited suggestions from the participants about what types of classes could benefit from the active code completion palettes. 119 participants suggested various types of classes and we classified the suggestions into several categories.

\subsubsection{Alternative/Tricky/Literal Syntax (16)}
These are the classes of which syntax is tricky, or complex. Even for a basic string class, people wanted to input multiline string or unicode string in more convenient way with automatic escaping. We address this issue in more detail in Section \#.
%
%\begin{itemize}
%	\item Example classes
%	
%	\begin{itemize}
%		\item String
%		\item Collections (e.g., dictionary)
%		\item Vector, Matrix
%		\item Embedded languages
%		\item URLs, Paths
%	\end{itemize}
%	
%	\item Expected features
%	
%	\begin{itemize}
%		\item Proper escaping
%		\item Syntax highlighting
%		\item Auto-completion
%	\end{itemize}
%\end{itemize}

\subsubsection{Unclear Parameter Implications (11)}
The classes in this category has some parameters in it, and it is not easy to predict the results of the parameter modifications. Instead of running the whole application to see the result, the palette can serve as a preview panel so that the developers may check the results directly. Also, the palette can be a control panel which facilitates modifying the parameters as in Juxtapose [?].

%\begin{itemize}
%	\item Example classes
%	
%	\begin{itemize}
%		\item Audio tweaks (modifying pitch, volume, etc.)
%		\item 3D transformation matrices
%		\item number / string / date formatting
%		\item message box / input box (e.g., \texttt{JOptionPane})
%	\end{itemize}
%\end{itemize}

\subsubsection{Query Languages (17)}
Similar to the regular expression and SQL palette we presented, the developers also wanted to check if the other types of query strings (e.g., XPath / XQuery) are correct or not by checking the results directly on the palette.

%\begin{itemize}
%	\item Example classes
%	
%	\begin{itemize}
%		\item Regular expression
%		\item SQL query
%		\item XPath / XQuery
%	\end{itemize}
%\end{itemize}

\subsubsection{Graphical Elements (27)}
The other popular category was graphical elements. We've already shown the color example, and the participants suggested many other graphical properties such as font, brush, shape and others. Also, people wanted to check or directly manipulate the GUI frames and layouts.

%\begin{itemize}
%	\item Example classes
%	
%	\begin{itemize}
%		\item Color, Shape, Brush, Font, etc.
%		\item JFrame, Layouts, any swing components
%	\end{itemize}
%\end{itemize}

\subsection{Complex Instantiation Procedures (5)}
Some of the participants pointed out that these palettes might be useful to create an object which requires complex instantiation code. For example, in order to read a text file in Java, the developer might want to use \texttt{BufferedReader} class. It is error prone to use this class because it requires try/catch block, and also one must close the file after reading it. By using a palette to choose a file or choose a variable which contains the file path, the developers could easily instantiate these objects. In the same way, it can alleviate the factory pattern usability problem \cite{ellis_factory_2007}. As long as the developers remember which class to use, they will not need to remember how to instantiate that class.

%\begin{verbatim}
%BufferedReader reader = null;
%try {
%    reader = new BufferedReader(
%        new FileReader(filePath));
%} catch (IOException e) {
%    e.printStackTrace();
%}
%// use reader here
%reader.close();
%\end{verbatim}

@@

%\begin{itemize}
%	\item Example classes
%	
%	\begin{itemize}
%		\item Keyboard Keys
%		\item Regular expression
%	\end{itemize}
%\end{itemize}

@@

>>>>

\usepackage{enumerate}
\usepackage{rotating}

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Usability Criteria for Graphical Code Completion}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Cyrus Omar}
\IEEEauthorblockA{Computer Science Department\\
Carnegie Mellon University\\
Pittsburgh, PA 15213\\
cyrus@cmu.edu}
\and
\IEEEauthorblockN{YoungSeok Yoon}
\IEEEauthorblockA{Institute for Software Research\\
Carnegie Mellon University\\
Pittsburgh, PA 15213\\
youngseok@cs.cmu.edu}
}

@@

% make the title area
\maketitle

%\begin{abstract}
%\boldmath
%\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
Many source code editors today feature interactive code completion. By invoking a keyboard shortcut, developers can summon a floating palette containing a list of contextually relevant variables, fields, methods, classes and code snippets. An adjacent floating window displays documentation for the currently highlighted item. This interface helps developers explore APIs and avoid basic spelling and logic errors [?], and empirical data indicates that professional developers invoke interactive code completion frequently \cite{murphy_how_2006}.

A number of new to the code completion list have been proposed in the literature. These methods combine information about the program context with make use of usage history [?], examples extracted from code repositories [?, ?] and crowdsourced information [?]

Calcite, reordering, crowdsourced code completion papers.

Our proposal: graphical code completion. Inversion of control: look at class. Also possible to look at external database. Visual languages have shown this might be useful in certain circumstances (?).

Consistent with best practices, we begin by seeking the usability criteria that govern whether such a system would be useful to professional developers. We conducted a large survey of these developers and extracted several criteria that both constrain the overall system architecture and inform the design (or caution against designing) palettes for different kinds of classes.

%\section{Tool}
%Described earlier.

\section{Methods}
Our subject pool consisted primarily of visitors to a large collaborative filtering website dedicated to programming called {\tt /r/programming} on reddit.com\footnote{http://www.reddit.com/r/programming}. At the time of our study, there were approximately 340,000 registered members. A small number of participants were also recruited using a mass email to local computer science graduate students. The recruitment materials in both cases stated that we were seeking developers ''familiar with an object-oriented programming language like Java, C\# or Visual Basic and an integrated development environment like Eclipse or Visual Studio''.

Participants were asked to take an online survey that would take approximately 20 minutes to complete. Of the 696 participants who started the survey, 475 participants finished it. Of these, 457 participants came from /r/programming and 18 from the graduate student mailing list. All quantitative metrics used in this study are based only on these participants.

\subsection{Familiarity with Languages and Editors}

We began by asking participants to rate their level of familiarity with several programming languages. Figure 1 summarizes these ratings. On average, participants rated themselves as very familiar with Java, C, C++ and JavaScript, familiar with C\#, Python and PHP and somewhat familiar with Visual Basic and Perl.

We also asked participants to select the integrated development environments and code editors that they were familiar with. The Eclipse IDE was familiar to 87.1\% of participants. This was followed by Visual Studio with 66.0\% of participants, Vi/Vim with 53.7\%, Netbeans with 37.7\%, Emacs with 24.8\% and IntelliJ IDEA with 16.4\%. Participants could also enter ``other'' choices. A number of editors and IDEs were entered, including Xcode, Textmate, Notepad++ and others.
%\begin{figure}
%  \includegraphics[width=3.5in,height=2in]{experience.png}
%  \caption{Participant familiarity with various programming languages. Starting at bottom: None, Somewhat familiar, Familiar, Very Familiar, Expert.}
%\end{figure}

\subsection{Mockups}
\begin{figure*}
  \includegraphics[scale=.6]{mockup-palette.png}
  \caption{The regular expression palette, shown in Eclipse.}
\end{figure*}
Next, we presented participants with a series of mockup palettes corresponding to a Color class, a regular expression Pattern class (Figure 2) and a SQL query ResultSet class. In addition to a screenshot of the palette itself, participants were shown mockups demonstrating how a user would invoke the palette, and a mockup showing the code that would be inserted once a selection had been made.

Before presenting each palette mockup we gathered information about the user's familiarity with that topic and the strategies that they would likely use to instantiate that class. The Color class was presumed familiar to all participants. Figure 3 summarizes responses for the regular expression and SQL classes. The low number of participants unfamiliar with these topics again provides evidence that our participants were not generally novice developers.

To determine the strategy typically used for instantiating the Color class, we asked the following question. The values in bold show the frequency of each response.

\begin{quote}
A user interface designer asks you to change the background color of a window to ``navy blue'' when the user presses a button. You have determined that you need to instantiate the 'Color' class to do so.
\\\\
Which of the following strategies would you use first to create an instance of the 'Color' class representing navy blue?\\

\begin{enumerate}[(a)]
\item I would use the IDE's content assist menu (see note below) to look for a predefined constant representing navy blue. (\textbf{58.4\%})
\item I would look up the documentation of the Color class to see if there is a predefined constant representing navy blue. (\textbf{19.0\%})
\item I would use an external tool (e.g. Photoshop or some other image editor) to determine the RGB values for navy blue. (\textbf{14.0\%})
\item I would guess at the RGB values for navy blue, then check whether my guess was correct by executing the application. (\textbf{5.3\%})
\item Other (please specify) (\textbf{3.3\%})\\
\end{enumerate}
\end{quote}

Figure 4 summarizes the responses to a more general question asked about regular expressions and SQL queries.

\begin{figure}
\begin{tabular}{lccc}
 & \textbf{Regular Expressions} &  & \textbf{SQL}\\
 \hline
Never used & 4.8\% &  & 0.0\%\\
Use infrequently & 46.7\% &  & 37.4\%\\
Use frequently & \textbf{48.4\%} &  & \textbf{62.6\%}\\
\hline
\end{tabular}
\caption{Participant's experience with regular expressions and SQL.}
\end{figure}

\begin{figure}
\begin{tabular}{lccc}
 & \textbf{Regular Expressions} &  & \textbf{SQL}\\
 \hline
Separate test script & 29.7\% &  & 15.5\%\\
Guess and check & 13.4\% &  & 16.0\%\\
External tool & \textbf{38.5\%} &  & \textbf{58.9\%}\\
Search for examples & 12.1\% & & 4.8\% \\
Other & 6.2\% & & 4.8\% \\
\hline
\end{tabular}
\caption{Typical strategies for regular expressions and SQL queries.}
\end{figure}

%For each of these classes,

%\subsubsection{Color}
%We began with a class representing colors. We asked participants to consider the following scenario:

\begin{figure}
\begin{tabular}{crccccc}\\\\
\textsc{class}
& 
& \begin{rotate}{20}Nearly every time\end{rotate}
& \begin{rotate}{20}Most of the time\end{rotate}
& \begin{rotate}{20}Some of the time\end{rotate}
& \begin{rotate}{20}Rarely\end{rotate}
& \begin{turn}{20}Never\end{turn}\\
\hline
Color &\vline& 9.6\% & 22.1\% & \textbf{32.4\%} & 28.2\% & 7.7\%\\
RegExp &\vline& \textbf{36.6\%} & 29.5\% & 21.8\% & 7.3\% & 4.8\%\\
SQL & \vline &18.2\% & 19.3\% & \textbf{30.9\%} & 20.4\% & 11.4\%\\
\hline
\end{tabular}
\caption{For each of the classes considered, participants were shown mockups of a code completion palette designed for instantiating that class, as well as the source code generated after parameters were entered. This table shows the distribution of responses to the question: ``Consider situations where you need to instantiate the [specified] class. What portion of the time, in these situations, do you think you would use this feature?''}
\end{figure}
%
%\begin{figure}
%  \includegraphics[scale=.6]{regex-insert.png}
%  \caption{Test}
%\end{figure}
%
%

%Then we asked:
%\begin{quote}
%\textbf{Q2} Consider situations where you need to instantiate the 'Color' class. What portion of the time, in these situations, do you think you would use this feature?
%\begin{enumerate}[(a)]
%\item Nearly every time (\textbf{9.6\%})
%\item Most of the time (\textbf{22.1\%})
%\item Some of the time (\textbf{32.4\%})
%\item Rarely (\textbf{28.2\%})
%\item Never (\textbf{7.7\%})\\
%\end{enumerate}
%\end{quote}
%

After presenting each palette, we asked users to rate the usefulness of each palette. Responses are summarized in Figure 5. We then allowed participants to provide qualifications and open-ended feedback by asking the following question: ``If you would like to qualify or elaborate on your answer to the previous question you may do so below. You may also suggest improvements for this palette.'' This information is analyzed in the remainder of the paper.

%\begin{quote}
%\textbf{Q4} How would you describe your level of familiarity with (regular expressions, SQL)?\\
%\\
%\begin{enumerate}[(a)]
%\item I do not know what [it is].	 (\textbf{0.7\%}, \textbf{0.0\%})
%\item I know what regular expressions are but I have not used them in practice. (\textbf{4.4\%, 0.0\%})
%\item I use regular expressions infrequently.	 (\textbf{46.0\%, 35.9\%})
%\item I use regular expressions frequently. (\textbf{49.0\%, 64.1\%})
%\end{enumerate}
%\end{quote}

%\begin{quote}
%\textbf{Q5} Consider situations where you need to write a regular expression to extract information from a string or collection of strings. What strategy would you most likely use first?

%\begin{enumerate}[(a)]
%\item I would write a separate test script to help come up with an appropriate regular expression. (\textbf{29.7\%})
%\item I would guess and check regular expressions by running my application. (\textbf{13.4\%})
%\item I would launch an external application or web tool to help me test regular expressions interactively. (\textbf{38.5\%})
%\item I would search the web for a sample regular expression suitable for my task. (\textbf{12.1\%})
%\item Other (please specify) (\textbf{6.2\%})
%\end{enumerate}
%\end{quote}

%\begin{quote}
%\textbf{Q6} Consider situations where you need to create a regular expression. What portion of the time, in these situations, do you think you would use this feature?
%\begin{enumerate}[(a)]
%\item Nearly every time (\textbf{36.6\%})
%\item Most of the time (\textbf{29.5\%})
%\item Some of the time (\textbf{21.8\%})
%\item Rarely (\textbf{7.3\%})
%\item Never (\textbf{4.8\%})\\
%\end{enumerate}
%\end{quote}

%\begin{quote}
%\textbf{Q7} How would you describe your level of familiarity with SQL?

%\begin{enumerate}[(a)]
%\item I do not know what regular expressions are.	 (\textbf{0.0\%})
%\item I know what SQL is I have not used it in practice. (\textbf{0.0\%})
%\item I use SQL infrequently.	 (\textbf{35.9\%})
%\item I use SQL frequently. (\textbf{64.1\%})\\
%\end{enumerate}
%\end{quote}

%\begin{quote}
%\textbf{Q8} Consider situations where you need to write a regular expression to extract information from a string or collection of strings. What strategy would you most likely use first?

%\begin{enumerate}[(a)]
%\item I would write a separate test script to help come up with an appropriate SQL query.	
 %(\textbf{15.5\%})
%\item I would guess and check by running my application. (\textbf{16.0\%})
%\item I would launch an external application or web tool to test my query against my database interactively.	(\textbf{58.9\%})
%\item I would search the web for an example SQL query suitable for my task.	
 %(\textbf{4.8\%})
%\item Other (please specify) (\textbf{4.8\%})
%\end{enumerate}
%\end{quote}

%Showed the SQL palette and asked:

%\begin{quote}
%\textbf{Q9} Consider situations where you need to create a SQL query. How often, in these situations, do you think you would use this feature?
%\end{quote}

\section{Design Implications}
\begin{figure}[b]
  \includegraphics[scale=.5]{color-palette.png}
  \caption{test}
\end{figure}

We next analyzed the open-ended responses to extract some design principles and constraints. We got 193 responses for the color palette, 129 for the regular expression palette, and 142 for the SQL palette. Also, we got total 119 suggestions about what other types of classes could benefit from similar types of palettes. The design can be divided into two main categories: system design constraints, and general palette design considerations. System design constraints are the criteria that should be taken into account in the design of the overall system architecture, and the general palette design considerations are about how to make more usable palettes.

@@

\subsubsection{Handling Separation of Concerns}

Professional developers were very wary of including constant data into the source code directly. Many responses expressed general sentiment or specific preference for project-wide color theme or external resource file. Also, many people stated that these tools might not be useful since they usually read an external resource file, or sometimes they explicitly suggested to add capability to handle the external resource files directly from the palette.

Examples of feedbacks in this category include:

\begin{itemize}
	\item The use of magic numbers in the source code is a bad design (54)
	\item Color picking is the `designer's job' (33)
	\item Would only use to define named constants (17)
	\item Would rather use color scheme / theme (3)
	\item Would use Object-Relational Mapping (ORM) / Language INtegrated Query (LINQ) instead of raw SQL query string (27 of 142)
	\item These palettes might be useful for prototyping, but not for real production (14)
	\item Instead of including the test cases into comments, they should be stored as unit tests (35 of 143)
\end{itemize}

Interestingly, most people seemed to consider regular expressions as part of the program logic. There were few people who complained about including the regular expression pattern string in the source code.

\subsubsection{Reversibility}

19 participants expressed that it should be able to bring the palettes back when they want to modify the parameters after they used the palettes.

\subsubsection{Palette Settings and State}

Many participants wanted to be able to configure the palettes or have the palettes maintain state even after the palettes are hidden. The followings are the comments that are related to this category.

\begin{itemize}
	\item Want recently used regular expressions, and control over the comments (12)
	\item Persistent database connection information is needed / want to manage connection pools (9)
	\item Want recent colors, capability of setting favorite colors (20)
\end{itemize}

\subsubsection{Interaction with Code Context}

Some people wanted to interact with the code context. For example, 13 people expressed that it would be great if the SQL query palette was capable of dynamically creating SQL query string by combining multiple string variables, or by assigning variables for the parameters.

\subsubsection{Performance}

@@

\subsubsection{IDE Independence}

@@

%\subsubsection{Composing palette logic}

\subsection{General Palette Design Considerations}
\subsubsection{Simplicity vs. Capability}

@@

\subsubsection{Keyboard Navigability}

@@

\subsubsection{Complexity}

%\subsubsection{Side Effects}

\section{Suggestions from the Participants}
We classified the suggestions into several categories. The number in the parentheses indicate the number of participants who suggested the classes which fall into the category.

\subsection{Alternative/Tricky/Literal Syntax (16)}
These are the classes of which syntax is tricky, or complex. Even for a basic string class, people wanted to input multiline string or unicode string in more convenient way with automatic escaping.

\begin{itemize}
	\item Example classes

	\begin{itemize}
		\item String
		\item Collections (e.g., dictionary)
		\item Vector, Matrix
		\item Embedded languages
		\item URLs, Paths
	\end{itemize}

	\item Expected features

	\begin{itemize}
		\item Proper escaping
		\item Syntax highlighting
		\item Auto-completion
	\end{itemize}
\end{itemize}

\subsection{Unclear Parameter Implications (11)}
The classes in this category has some parameters in it, and it is not easy to predict the results of the parameter modifications. Instead of running the whole application to see the result, the palette can serve as a preview panel so that the developers may check the results directly. Also, the palette can be a control panel which facilitates modifying the parameters.
\begin{itemize}
	\item Example classes

	\begin{itemize}
		\item Audio tweaks (modifying pitch, volume, etc.)
		\item 3D transformation matrices
		\item number / string / date formatting
		\item message box / input box (e.g., \texttt{JOptionPane})
	\end{itemize}
\end{itemize}

\subsection{Query Languages (17)}
Similar to the unclear parameter implications category, the developers wanted to check if the query string is correct or not by checking the results directly on the palette.

\begin{itemize}
	\item Example classes

	\begin{itemize}
		\item Regular expression
		\item SQL query
		\item XPath / XQuery
	\end{itemize}
\end{itemize}

\subsection{Graphical Elements (27)}
The other popular category was graphical elements. We've already shown the color example, and the participants suggested many other graphical properties such as font, brush, shape and others. Also, people wanted to check or directly manipulate the GUI frames and layouts.

\begin{itemize}
	\item Example classes

	\begin{itemize}
		\item Color, Shape, Brush, Font, etc.
		\item JFrame, Layouts, any swing components
	\end{itemize}
\end{itemize}

\subsection{Complex Instantiation Procedures (5)}
Some of the participants pointed out that these palettes might be useful to create an object which requires complex instantiation code. For example, in order to read a text file in Java, the developer might want to use \texttt{BufferedReader} class. The instantiation code might look as following:
\begin{verbatim}
BufferedReader reader = null;
try {
    reader = new BufferedReader(
        new FileReader(filePath));
} catch (IOException e) {
    e.printStackTrace();
}
// use reader here
reader.close();
\end{verbatim}
By using a palette to choose a file or choose a variable which contains the file path, the developers could easily instantiate these objects. In the same way, it can alleviate the factory pattern usability problem. As long as the developers remember which class to use, they will not need to remember how to instantiate that class.

\begin{itemize}
	\item Example classes

	\begin{itemize}
		\item BufferedReader
		\item Classes using factory patterns
		\item Database connection string builder
	\end{itemize}
\end{itemize}

@@

\begin{itemize}
	\item Example classes

	\begin{itemize}
		\item Keyboard Keys
		\item Regular expression
	\end{itemize}
\end{itemize}

@@

%\begin{itemize}
%\item 54 expressed a concern about encouraging inlining of constants, preferring external resource files. 17 stated that they would use this feature only to define constants. 33 stated that it should be the designer's job.
%\item 11 expressed a concern about keyboard navigability
%\item 27 expressed a concern about the complexity of the palette
%\item 10 expressed an interest in a color picker. 7 wanted a multitabbed window with different options for picking colors.
%\item 26 expressed a desire for color names, 10 of which wanted to ensure color constant names were used if available
%\item 20 wanted persistent state (MRU, favorites)
%\end{itemize}

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

\section{Conclusion}
Overall, professional developers appear willing to use these for a number of use cases. However, the usability criteria described above impose stiff requirements on the architecture. Of particular concern, we found that the developers preferred help with program logic, but not the constant data. This indicates that another area of application may be in specialized editors for resource files.
